---
title: "Durbin-Levinson Recursion for an MA1 Model"
output: md_document
---

```{r}
library("stats")
```

### Model Setup

Specify the MA(1) model parameter $\theta = 0.8$ and the simulation length $n = 200$.

```{r}
theta <- 0.8

n <- 200
```

### Simulate MA(1) Process

The function `simulate_ma1` generates a realization of the MA(1) model
$$
X_t = W_t + \theta W_{t-1},
$$
where $W_t$ is standard normal noise.  The function returns a numeric vector of length $n$.

```{r}
simulate_ma1 <- function(n, theta) {
    
    w <- rnorm(n)
    
    x <- w + theta * c(0, w[-n])
    
    return(x)
    
}
```

The function is used to generate two independent datasets.

### Durbin-Levinson Recursion

The function `durlev_ma1` computes the lag-2 Durbin-Levinson coefficients for the MA(1) model.  It first computes the autocovariances
$$
\gamma(0) = 1 + \theta^2, \quad \gamma(1) = \theta, \quad \gamma(2) = 0,
$$
then applies the recursion to obtain $\phi_{1,1}, \phi_{2,1}, \phi_{2,2}$.  These coefficients depend only on $\theta$, not on the simulated data.  The function prints a labeled block and returns the coefficients.

```{r}
durlev_ma1 <- function(tag, theta) {

    gamma_0 <- 1 + theta^2
    gamma_1 <- theta
    gamma_2 <- 0

    phi_11 <- gamma_1 / gamma_0
    phi_22 <- -phi_11^2
    phi_21 <- phi_11 * (1 - phi_11^2)

    coeffs <- list(gamma = c(g0 = gamma_0, 
                             g1 = gamma_1, 
                             g2 = gamma_2),
                   phi = c(phi_11 = phi_11, 
                           phi_21 = phi_21, 
                           phi_22 = phi_22))

    cat("D-L Coefficients for", tag, "\n"); print(coeffs)

    return(coeffs)
    
}
```

This function is called separately for each dataset.

### Generate Datasets

I generate two datasets using different seeds so the simulated values differ.

```{r}
set.seed(123)
xA <- simulate_ma1(n, theta)

set.seed(456)
xB <- simulate_ma1(n, theta)
```

These datasets will produce different predictions and recoveries.

### Compute Coefficients

I compute the recursion coefficients for each dataset.  Even though the labels differ, the coefficients themselves are identical because they depend on on $\theta$.

This code prints two labeled blocks showing the same coefficient values.

```{r}
coeffsA <- durlev_ma1("xA (Seed 123)", theta)

coeffsB <- durlev_ma1("xB (Seed 456)", theta)
```

### One-Step-Ahead Prediction

The one-step-ahead predictor uses
$$
\hat{X}_3 = \phi_{2,1} X_2 + \phi_{2,2} X_1.
$$

Because $X_1$ and $X_2$ differ between `xA` and `xB`, the predictions differ.  This section demonstrates the data‑dependent nature of the predictor.

```{r}
predict_X3 <- function(x, coeffs) {
    
    phi_21 <- coeffs$phi["phi_21"]
    
    phi_22 <- coeffs$phi["phi_22"]
    
    phi_21 * x[2] + phi_22 * x[1]
    
}

predA_X3 <- predict_X3(xA, coeffsA)

predB_X3 <- predict_X3(xB, coeffsB)

cat("Prediction of X3 (Seed 123)\n"); predA_X3

cat("Prediction of X3 (Seed 456)\n"); predB_X3

```

### One-Step-Behind Recovery

Again, the coefficients are identical, but the recovered values differ because the data differ.

```{r}
recover_X3 <- function(x, coeffs) {
    
    phi_21 <- coeffs$phi["phi_21"]
    
    phi_22 <- coeffs$phi["phi_22"]
    
    phi_21 * x[5] + phi_22 * x[4]
    
}

recA_X3 <- recover_X3(xA, coeffsA)

recB_X3 <- recover_X3(xB, coeffsB)

cat("Recovery of X3 (Seed 123)\n"); recA_X3

cat("Recovery of X3 (Seed 456)\n"); recB_X3
```

This confirms that Durbin-Levinson coefficients are model-fixed, while predictions and recoveries are data-dependent.
