---
title: "W5 MGA Part 3 Extension"
output: md_document
---

```{r}
library("knitr")
```

I define a function that computes the components of the sample autocorrelation at lag 1 with the sample size $n$, the numerator

$$
\sum^n_{t=2} (x_t - \bar{x})(x_{t-1} - \bar{x}),
$$

the denominator
$$
\sum^n_{t=1} (x_t - \bar{x})^2,
$$

and the sample ACF
$$
\rho_1 = \frac{numerator}{denominator}
$$

```{r}
acf_components <- function(x) {
    
    n <- length(x)
    
    xbar <- mean(x)
    
    num <- sum((x[2:n] - xbar) * (x[1:(n-1)] - xbar))
    
    denom <- sum((x - xbar)^2)
    
    rho1 <- num / denom
    
    list(n = n,
         numerator = num,
         denominator = denom,
         acf1 = rho1)
    
}
```

I apply the function to four datasets of increasing size.

```{r}
x_small <- c(2, 3, 1, 4, 2)

res_small <- acf_components(x_small)

x_large <- c(2, 3, 1, 4, 2, 3, 2, 1, 4, 3)

res_large <- acf_components(x_large)

x_20 <- c(2,3,1,4,2,3,2,1,4,3,2,3,1,4,2,3,2,1,4,3)

res_20 <- acf_components(x_20)

x_50 <- rep(c(2,3,1,4,2,3,2,1,4,3), 5)

res_50 <- acf_components(x_50)

x_100 <- rep(c(2,3,1,4,2,3,2,1,4,3), 10)

res_100 <- acf_components(x_100)
```

And present the results in table form.

```{r}
results_df <- data.frame(
    
    sample = c("x_small", 
               "x_large", 
               "x_20", 
               "x_50",
               "x_100"),
    
    n = c(res_small$n, 
          res_large$n, 
          res_20$n, 
          res_50$n,
          res_100$n),
    
    numerator = c(res_small$numerator, 
                  res_large$numerator, 
                  res_20$numerator, 
                  res_50$numerator,
                  res_100$numerator),
    
    denominator = c(res_small$denominator, 
                    res_large$denominator, 
                    res_20$denominator, 
                    res_50$denominator,
                    res_100$denominator),
    
    acf1 = c(res_small$acf1, 
             res_large$acf1, 
             res_20$acf1, 
             res_50$acf1,
             res_100$acf1)
    
)

kable(results_df, caption = "ACF Components at Lag 1 for Each Sample")
```

As the sample grows, the estimates converge to $-0.52$.  The table shows the expected behavior of the autocorrelation estimator.  Small samples produce volatile ACF estimates because the numerator and denominator are sensitive to individual observations.  Larger samples produce more consistent ACF estimates and converge to the true autocorrelation.  The lag-1 estimator becomes more reliable as $n$ increases.


